---
title: "How are the penguins going?"
author: "Yan Holtz"
date: today
format:
  html:
    toc: true
    css: style.css
    code-fold: true
    code-summary: "Show the code"
---

## Introduction

This document analyses what's into the famous Palmer Penguins dataset. The dataset has been found [on the web](https://gist.github.com/slopp/ce3b90b9168f2f921784de84fa445651) and copied into this repository..

## Libraries

Analyzing data with R always requires a few libraries. Let's load them all at the top of this document. If you are trying to replicate this work, you probably want to install them first with `install.packages()`\`

```{r, warning=FALSE, message=FALSE}
library(ggplot2)  # to create most of the charts
library(DT)       # to show an interactive table in this doc
library(hrbrthemes)
```

## Data loading

Let's load it!

```{r}
# Load data --------------------------------------
data <- read.csv("data/penguins.csv")

```

We now have a tibble (= a data frame) that we can start exploring. In case you are curious, this is how the dataset looks like:

```{r}
# Show data --------------------------------------
datatable(data,
  rownames = FALSE,
  filter = "top", options = list(pageLength = 5, scrollX = T)
)

```

## Relationship between **Bill length vs. depth**

Let's start by checking the relationship between Bill Length and Bill depth. At first sight, it looks like there is an **inverse correlation** between those 2 variables:

```{r, warning=FALSE, message=FALSE}
ggplot(
  data,
  aes(
    x = bill_length_mm,
    y = bill_depth_mm
  )
) +
  geom_point() +
  labs(
    title = "Penguin bill dimensions",
    subtitle = "Palmer Station LTER",
    x = "Bill length (mm)",
    y = "Bill depth (mm)"
  ) +
  theme(
    plot.title.position = "plot",
    plot.caption = element_text(hjust = 0, face = "italic"),
    plot.caption.position = "plot"
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "gray50") +
  theme_ipsum()

```

This result being surprising, it is possibly a great idea to check the result specie per specie, in the quest for a Simpson Paradox:

```{r, warning=FALSE, message=FALSE}
ggplot(
  data,
  aes(
    x = bill_length_mm,
    y = bill_depth_mm,
    group = species
  )
) +
  geom_point(
    aes(
      color = species,
      shape = species
    ),
    size = 3,
    alpha = 0.8
  ) +
  geom_smooth(method = "lm", se = FALSE, aes(color = species)) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    title = "Penguin bill dimensions",
    subtitle = "Bill length and depth for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Penguin species",
    shape = "Penguin species"
  ) +
  theme_ipsum()
  

```
